Discrete Quantum Theories and Computing

### 2018-12-13 dissertation.tex
dissertation.lyx "Chapter 1 Introduction" Updating...

Since we as human cannot distinguish quantities differing only by
an arbitrarily small extent \cite{Turing_1937,Gisin2017}, quantum
mechanics and computation based on the continuum of complex numbers
may not capture our ability of predicting quantum phenomenon and the
power of quantum computation. Being agnostic about whether the reality
is ultimately discrete or continuous, we considered two types of quantum
models to incorporate our limitation of distinguishability. By assuming
the quantum states are ultimately quantized and distinguishable, our
first type of quantum theories replaces the complex numbers by discrete
finite fields. We then focus on the limitation of distinguishability
itself, incorporate the finite precision measurement into our model,
and consider quantum interval-valued probability measures.

After first suggesting by Richard Feynman 40 years ago~\cite{Feynman1982Simulating},
IBM, \textitalian{Rigetti}, Google, Microsoft, and many other companies
and countries start to build realistic quantum computers \cite{Shieber2018,Kahn2019},
and seek more efficient ways to simulate chemical molecules and to
break down RSA \cite{boneh1999twenty,wiki:RSAFactoring}. Despite
how we compute these two tasks, we can notice they require us to compute
two different types of quantities. While RSA requires discrete computation,
it is beyond our ability to distinguish whether the quantities required
by simulating molecules are discrete or not. Before discussing the
quantities characterizing and computing with quantum mechanics, it
is instructive to review how classical physical quantities are characterized
and computed.

Discrete quantities can be easily characterized by whole numbers~\cite{ParkerBaldridge2004}.
Even if the proportion of whole numbers are fractions, the operations
among fractions is essentially the same as those among whole numbers
over the common denominator, and the computational results faithfully
represent the corresponding quantities. In general, discrete computation
can be reliably repeated with exact equivalence, and any digital computer
can faithfully simulate a Turing machine limited only by the available
memory and time~\cite{Turing_1937}.

In contrast, the non-apparently discrete quantities confused many
cultures since very beginning. For example, Greek mathematicians in
the school of Pythagoras tried to apply their discrete idea into their
model, and believed every mathematical model must be ultimately discrete.
They believed so much that they in legend murdered Hippasus because
he found there is no minimum unit between a side of a square and its
diagonal. Comparing the idea that the length of a square's diagonal
is not a physical quantity or there is no square physically, working
with a mathematical model without minimum unit seems more appealing,
and their corresponding physical quantities are also assumed to be
continuous. An ancient Chinese philosopher Chuang-tzu (庄子) expressed
this belief as ``If a one-foot-long stick is cut into halves every
day, the cutting will never come to an end''\footnote{``一尺之棰，日取其半，萬世不竭。'' in Chinese \cite{Zhuangzi1999,Liu2018}.}.
Later, these continuous quantities becomes the basic component of
analog computers. After logarithm was discovered, a slide rule, also
known as a slipstick, is used to compute multiplication, division,
and more complex operations as a mechanical analog computer~\cite{wiki:SlideRule}.
Charge amplifiers is used to build integrators \cite{wiki:Integrator,wiki:ChargeAmplifier}.

>>>>

Since we as human cannot distinguish quantities differing only by
an arbitrarily small extent \cite{Turing_1937,Gisin2017}, the quantum
mechanical model based on the continuum of complex numbers may not
capture our ability of predicting quantum phenomenon and the power
of quantum computation. Being agnostic about whether the reality is
ultimately discrete or continuous, we considered two types of quantum
models to incorporate our limitation of distinguishability. By assuming
the quantum states are ultimately quantized and distinguishable, our
first type of quantum theories replaces the complex numbers by discrete
finite fields. We then focus on the limitation of distinguishability
itself, incorporate the finite precision measurement into our model,
and consider quantum interval-valued probability measures.

After first suggesting by Richard Feynman 40 years ago~\cite{Feynman1982Simulating},
IBM, \textitalian{Rigetti}, Google, Microsoft, and many other companies
and countries start to build realistic quantum computers recently
\cite{Shieber2018,Kahn2019}, and seek more efficient ways to simulate
chemical molecules and to break down RSA \cite{boneh1999twenty,wiki:RSAFactoring}.
No matter whether we use classical or quantum computers to solve these
two tasks, we notice that they require two different types of quantities.
The quantities required by RSA are discrete, and it is beyond our
ability to distinguish whether the quantities required by simulating
molecules are discrete or not. Before we dive into the discussion
of these quantities in quantum computation, it is instructive to review
how classical quantities are characterized and computed.

Discrete quantities can be easily characterized by whole numbers~\cite{ParkerBaldridge2004}.
Even if the proportion of whole numbers are fractions, the operations
among fractions are essentially the same as those among whole numbers
over the common denominator. In general, discrete computation can
be reliably repeated with exact equivalence, and any digital computer
can faithfully simulate a Turing machine limited only by the available
memory and time~\cite{Turing_1937}.

In contrast, the non-apparently discrete quantities confused many
cultures since very beginning. On one hand, the ancient Chinese philosopher
Chuang-tzu (庄子) didn't believe these quantities have a minimum unit,
and expressed ``If a one-foot-long stick is cut into halves every
day, the cutting will never come to an end''\footnote{``一尺之棰，日取其半，萬世不竭。'' in Chinese \cite{Zhuangzi1999,Liu2018}.}.
On the other hand, Greek mathematicians in the school of Pythagoras
believed every mathematical model must be ultimately discrete. They
believed so much that they in legend murdered Hippasus because he
found there is no minimum unit between a side of a square and its
diagonal. Comparing the idea that the length of a square's diagonal
is not a physical quantity or there is no square physically, working
with a mathematical model without minimum unit seems more appealing.
Since the mathematical model is continuous, their corresponding physical
quantities are also assumed to be continuous. Later, these continuous
quantities becomes the basic component of analog computers. After
logarithm was discovered, a slide rule, also known as a slipstick,
is used to compute multiplication, division, and more complex operations
as a mechanical analog computer~\cite{wiki:SlideRule}. Charge amplifiers
is used to build electrical analog computers and compute the integration
in calculus \cite{wiki:Integrator,wiki:ChargeAmplifier}.
